---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r load_data, }
        library(ggplot2)
        data <-read.csv('./activity.csv')
        #clean data
        data_1 <- na.omit(data)
```
###Histogram
Plot a histogram to see the number of steps that appears more oftern per day.
```{r plot_histogram, }
        ggplot(data_1, aes(x=steps, fill=date)) + 
                geom_histogram(binwidth=80) +
                ggtitle("Total number of steps taken each day")+
                scale_fill_grey()

```

Code to generate the mean and median of the data
```{r, mean_median_total_Steps, results="asis"}
        library(xtable)
        #split the data frame
        split_data <- split(data_1, data_1$date)
        #total number of steps per day
        sum_step=sapply(1:61, function(x) sum(unlist(split_data[[x]][1])))
        #find summary of the steps per day
        size<-length(split_data)
        new_data <- data.frame(Id=1:61,
                        median=sapply(1:61, function(x) median(unlist(split_data[[x]][1]))), 
                        mean=sapply(1:61, function(x) mean(unlist(split_data[[x]][1]))))
        xt<-xtable(new_data)
        print(xt,type="html")
```
## What is mean total number of steps taken per day?
```{r, steps_mean_total, results="asis"}
split_data <- split(data$steps,data$date)
        new_data <- as.data.frame(split_data)
        #find the avereage accross all days
        step_mean<-rowMeans(new_data,na.rm = TRUE)
        #set interval
        split_interval <- split(data$interval,data$date)
        df <- data.frame(interval=split_interval[[1]], means_step=step_mean)

        #find the max avg step
        max_mean_step <- sort(df$means_step, decreasing=TRUE)[1]
        #Print the max
        question_2<-df[df$means_step==max_mean_step,]
        xt<-xtable(question_2)
        print(xt,type="html")
```

## What is the average daily activity pattern?
The next graph shows the average daily activity pattern
```{r, time_series_avg_steps}
        ggplot( data = df, aes( means_step, interval )) + 
                geom_line() +
                ggtitle("Time Series of steps (Avg per month )")+
                xlab("Time Series - Month") +
                ylab("Average of steps") 
```
## Imputing missing values
Find the missing values
```{r, find_missing_values}
na_data<- apply(data, 1, function(x){any(is.na(x))})
        na_data<-data[na_data,]
```

Generate new data frame for the NA values. Replace the values with the mean of the interval.
```{r, generate_new_data_frame_na}
split_intevals <- split(data$interval,data$date)
        dfM <- data.frame(step_mean,interval=split_intevals[1])
        update_values <- function(step,interval) if(is.na(step)) { step=dfM[dfM$X2012.10.01==interval,][1,1] }else{ step }
        new_df <-data
        new_df$steps <- mapply(update_values,new_df$steps,new_df$interval)
```

Generate new histogram plot
```{r plot_histogram_without_na }
        ggplot(new_df, aes(x=steps, fill=date)) + 
                geom_histogram(binwidth=80) +
                ggtitle("Total number of steps taken each day without NAs")+
                scale_fill_grey()
```

Generate mean and median
```{r, mean_median_without_na, results="asis"}

        #split the data frame
        split_data_na <- split(new_df, new_df$date)
        #total number of steps per day
        sum_step=sapply(1:61, function(x) sum(unlist(split_data_na[[x]][1])))
        #find summary of the steps per day
        new_data_na <- data.frame(Id=1:61,
                        median=sapply(1:61, function(x) median(unlist(split_data_na[[x]][1]))), 
                        mean=sapply(1:61, function(x) mean(unlist(split_data_na[[x]][1]))))
        xt<-xtable(new_data_na)
        print(xt,type="html")
```
## Are there differences in activity patterns between weekdays and weekends?
Create the factors
```{r, create_Factors, results="asis"}
df_last<-data
        df_last$day <- weekdays(as.Date(df_last$date))
        v_weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
                        "Friday")
        
        v_weekend <- c("Saturday","Sunday")
        
        df_weekdays <- df_last[df_last$day %in% v_weekdays ,]
        df_weekend <- df_last[df_last$day %in% v_weekend ,]
        
        #split the data for week days
        split_data <- split(df_weekdays$steps,df_weekdays$day)
        new_data <- as.data.frame(split_data)
        #find the avereage accross all days
        step_mean<-rowMeans(new_data,na.rm = TRUE)
        #set interval
        split_interval <- split(df_weekdays$interval,df_weekdays$date)
        df1 <- data.frame(interval=split_interval[[1]], means_step=step_mean)
        
        #find the max avg step
        max_mean_step <- sort(df1$means_step, decreasing=TRUE)[1]
        #Print the max
        df1[df1$means_step==max_mean_step,]
        
        split_data <- split(df_weekend$steps,df_weekend$day)
        new_data <- as.data.frame(split_data)
        step_mean<-rowMeans(new_data,na.rm = TRUE)
        df2 <- data.frame(interval=split_interval[[1]], means_step=step_mean)
```
Plot the data
```{r, plot_facotrs, fig.keep ='all' }
        #plot
        ggplot( data = df2, aes( means_step, interval )) + 
                geom_line() +
                ggtitle("Weekend")+
                xlab("Time Series - Month") +
                ylab("Average of steps") 
        
        ggplot( data = df1, aes( means_step, interval )) + 
                geom_line() +
                ggtitle("Week")+
                xlab("Time Series - Month") +
                ylab("Average of steps")   
```
